---
title: "NZ Cohort Life Tables"
author: "Duncan Garmonsway"
date: "`r Sys.Date()`"
output: rmarkdown::github_document()
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `nzlifetables` package distributes the Statistics New Zealand Cohort Life
Tables
[spreadsheet](http://www.stats.govt.nz/browse_for_stats/health/life_expectancy/cohort-life-tables-info-releases.aspx)
in convenient, `tidy` form.

```{r, echo = FALSE}
suppressPackageStartupMessages(library(tidyverse))
library(nzlifetables)
nzlifetables %>%
  mutate(year = cohort + age,
         year_of_death = year + ex) %>%
  ggplot(aes(year, year_of_death, colour = cohort,
             group = interaction(cohort, projected),
             alpha = -projected)) +
  geom_line() +
  scale_x_continuous(breaks = full_seq(c(1880, 2040), 20)) +
  scale_colour_viridis_c() +
  scale_alpha(range = c(.3, 1), guide = FALSE) +
  facet_wrap(~ sex, nrow = 1) +
  ylab("Estimated year of death") +
  ggtitle("Cohort life expectancy",
          sub = "Estimated year of death, given that a person is still alive in a given year")
```

## Variables

* `sex`
* `cohort`
* `age` Exact age (years)
* `lx` Number alive at exact age, out of 100,000 born
* `Lx` Average number alive in the age interval, out of 100,000 born
* `dx` Number dying in the age interval, out of 100,000 born
* `px` Probability that a person who reaches this age lives another year
* `qx` Probability that a person who reaches this age dies within a year
* `mx` Central death rate for the age interval
* `sx` Proportion of age group surviving another year
* `ex` Expected number of years of life remaining at age x
* `projected` Whether or not the statistic is a projection

## Installation

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("nacnudus/nzlifetables")
```

## Examples

```{r}
library(tidyverse)
library(nzlifetables)
```

```{r}
nzlifetables %>%
  mutate(year = cohort + age) %>%
  ggplot(aes(year, ex, colour = age,
             group = interaction(age, projected),
             alpha = -projected)) +
  geom_line() +
  scale_colour_viridis_c() +
  scale_alpha(range = c(.3, 1), guide = FALSE) +
  facet_wrap(~ sex) +
  ggtitle("Cohort life expectancy",
          sub = "Remaining lifespan, given age in a given year")
```

```{r}
nzlifetables %>%
  mutate(year = cohort + age) %>%
  ggplot(aes(year, lx, colour = age,
             group = interaction(age, projected),
             alpha = -projected)) +
  geom_line() +
  scale_colour_viridis_c() +
  scale_alpha(range = c(.3, 1), guide = FALSE) +
  facet_wrap(~ sex) +
  ggtitle("Cohort life expectancy",
          sub = "Number alive at exact age in a given year, out of 100,000")
```

```{r}
nzlifetables %>%
  mutate(year = cohort + age,
         year_of_death = year + ex) %>%
  ggplot(aes(year, year_of_death, colour = cohort,
             group = interaction(cohort, projected),
             alpha = -projected)) +
  geom_line() +
  scale_x_continuous(breaks = full_seq(c(1880, 2040), 20)) +
  scale_colour_viridis_c() +
  scale_alpha(range = c(.3, 1), guide = FALSE) +
  facet_wrap(~ sex, nrow = 1) +
  ylab("Estimated year of death") +
  ggtitle("Cohort life expectancy",
          sub = "Estimated year of death, given that a person is still alive in a given year")
```

```{r}
nzlifetables %>%
  mutate(year = cohort + age,
         lifespan = age + ex) %>%
  ggplot(aes(year, lifespan, colour = cohort,
             group = interaction(cohort, projected),
             alpha = -projected)) +
  geom_line() +
  scale_x_continuous(breaks = full_seq(c(1880, 2040), 20)) +
  scale_colour_viridis_c() +
  scale_alpha(range = c(.3, 1), guide = FALSE) +
  facet_wrap(~ sex, nrow = 1) +
  ylab("Estimated lifespan") +
  ggtitle("Cohort life expectancy",
          sub = "Estimated lifespan, given that a person is still alive in a given year")
```

```{r}
nzlifetables %>%
  mutate(lifespan = age + ex) %>%
  ggplot(aes(age, lifespan, colour = cohort,
             group = interaction(cohort, projected),
             alpha = -projected)) +
  geom_line() +
  geom_abline(intercept = 0, slope = 1) +
  scale_x_continuous(breaks = full_seq(c(0, 100), 10)) +
  scale_colour_viridis_c() +
  scale_alpha(range = c(.3, 1), guide = FALSE) +
  facet_wrap(~ sex, nrow = 1) +
  ylab("Estimated lifespan") +
  ggtitle("Cohort life expectancy",
          sub = "Estimated lifespan, given that a person is still alive at a given age")
```

```{r}
nzlifetables %>%
  ggplot(aes(age, qx, colour = cohort,
             group = interaction(cohort, projected),
             alpha = -projected)) +
  geom_line() +
  scale_x_continuous(breaks = full_seq(c(0, 100), 10)) +
  scale_colour_viridis_c() +
  scale_alpha(range = c(.3, 1), guide = FALSE) +
  facet_wrap(~ sex, nrow = 1) +
  ylab("Probabilitiy of dying with a year") +
  ggtitle("Cohort life expectancy",
          sub = "Probability of dying, given that a person is still alive at a given age")
```

```{r}
nzlifetables %>%
  ggplot(aes(age, lx, colour = cohort,
             group = interaction(cohort, projected),
             alpha = -projected)) +
  geom_line() +
  scale_x_continuous(breaks = full_seq(c(0, 100), 10)) +
  scale_colour_viridis_c() +
  scale_alpha(range = c(.3, 1), guide = FALSE) +
  facet_wrap(~ sex, nrow = 1) +
  ylab("Number still alive at given age") +
  ggtitle("Cohort life expectancy",
          sub = "Number still alive at a given age, out of 100,000")
```
